Robust Spectroscopic Inference with Imperfect Models

### 2014-11-02 tex/paper.tex
rough draft of appendix complete

The procedure outlined above can be a significant computational challenge.  A typical data spectrum 
has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the many repeated evaluations of the matrix 
product $\vR^{\trans} \vC^{-1} \vR$ in the likelihood calculation can be numerically expensive; 
prohibitively so in the brute force sense, since $\vC$ is non-trivial by design.  However, there 
are two aspects of these calculations that enable us to take advantage of some clever algorithms 
designed for such issues.  First, since we are only interested in the product $\vR^{\trans} 
\vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can employ some efficient sparse matrix approximations 
that avoid the computational expense of matrix inversion.  And second, because $\vC$ is a 
symmetric, positive semi-definite matrix, we can employ Cholesky factorization to optimize the 
evaluation of the matrix product.  In practice, this means Steps~(2), (3a), and (3b) in the 
procedure outlined above can be conducted with minimal computation cost.  Although the Cholesky 
factorization must be re-done for each update of $\vp_{\mathsf C}$ (Step~3d), this does not incur a 
high cost because the covariance kernels were designed to deliver a sparse matrix (via the Hann 
windows).  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization 
operations \citep{chen08,davis09}, and extend the Metropolis-Hastings sampler included in the {\tt 
emcee} package \citep{foreman-mackey13} to function within a blocked Gibbs 
sampler.\footnote{\url{https://github.com/iancze/emcee}} For echelle spectra with multiple orders or multiple spectra of the same object, taken with different instruments, the nuisance parameters for each chunk of spectrum are independent. Additionally, the computationally intensive steps of generating a model spectrum for a specific wavelength range and evaluating the likelihood on this spectral chunk can also be parallelized. Therefore, the only part of the code that needs to be synchronized is the MCMC proposal of stellar parameters. Because all of the computationally expensive steps can be parallelize, this algorithim has the ability to be massively parallelized on a cluster, enabling simultaneous inference for wide spectral ranges or multiple spectral datasets acquired with different instruments.\\

@@

Alternatively, this kind of feedback could be used to make data-driven modifications to the already 
existing models, creating a new library of semi-empirical spectral models.  This could be 
accomplished by linking the parameters of the covariance kernels derived from fitting many spectra 
in a hierarchical Bayesian model, which would add confidence to the assessment that certain 
spectral features are {\it systematic} outliers and offer general quantitative guidance on how to 
weight them in the likelihood calculation.  Rather than simply assembling an empirical spectral 
library using only observations, this combined machine-learning approach would naturally provide a 
physical anchoring for the key physical parameters, since they are reflected in the spectra based 
on the physical assumptions in the original models.  This kind of large-scale analysis holds great 
promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
like those being collected in programs like the APOGEE and HERMES surveys), since they provide 
enormous leverage for identifying and improving the underlying model systematics. \\
%Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
%include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
%linked to that, since there is a popular Herschel HERMES program that will confuse people...

@@

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}. We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}. Prior to the PCA decomposition, we first identify a subset of $M$ synthetic spectra which will be relevant to the star at hand. We then standardize the synthetic library by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid. The mean spectrum is 
\begin{equation}
  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
\end{equation}
$\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of stellar parameters sampled in the spectral library, while $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection. All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
\begin{equation}
  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are computed from the standardize grid using PCA analysis (\cite{ivezic13}). We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$). We denote a single $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$. As an example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.

@@

We model each eigenspectrum weight independently; therefore there is a separate Gaussian process ($\mathcal{GP}$) for each eigenspectrum. For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$. The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
\begin{equation}
  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( 0, \mathcal{K}_\textrm{SE} (\{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})\right )
\end{equation}
The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$. Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
\begin{widetext}
\begin{equation}
  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2 + (\log g_i - \log g_j)^2 + (\Z_i - \Z_j)^2 }{2 l_\textrm{SE}^2} \right )
\end{equation}
\end{widetext}
This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals. Then the $M$ by $M$ covariance matrix is 
\begin{equation}
\mathsf{C}^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | a_\textrm{SE}, l_\textrm{SE})
\end{equation}
which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters.

@@

Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
\begin{equation}
  \begin{bmatrix}
    \mathbf{w}_k^\textrm{grid}\\
    w_k
  \end{bmatrix} \sim
  \mathcal{N} \left (
  \mathbf{0},
  \begin{bmatrix}
    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
\mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
  \end{bmatrix}
  \right )
\end{equation}
where the dependence of the covariance kernel on the optimized hyperparameters is implied. Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$. And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$. The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$). The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
\begin{multline}
  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) = \\
  \mathcal{N} \Bigl ( w_k \Bigl |\; \Bigr . \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}, \Bigr . \\
 \Bigl . \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \Bigr )
 \label{eqn:conditional_gp}
\end{multline}
Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).

@@

@@

EDIT As an aside, this process of prediction is completely analogous to the process by which we generated the residual draws in the the covariance diagram. These were draws from the prior distribution (specified by Cdata). This is similar to drawing random values from the covariance matrix in Section XX given covariance matrices filled in by covariance kernels. Drawing from one giant Gaussian. Only now, we are drawing from a giant Gaussian conditioned on several points, which are the weights at the grid point locations. The benefit of using a Gaussian process for prediction/regression is that we now specified a probability distribution over possible interpolated values of the weights. We can draw multivariate samples from this probability distribution, like we did for the covariance matrices in Section XX to demonstrate the residuals (add note there to see here), only now we can also condition on the observed weight values. Notice how the GP directly interpolates the values at the grid points. For example, Fig A.2 (middle) shows draws from vectors of densely packed stellar values in order to show that simultaneous multivariate draws along a vector exhibit a certain structure of correlations (similar to the residuals in Section XX. However, the spread of subsequent multivariate draws at a specific single stellar value is given by a Gaussian distribution (Fig A.2 right).

So far, we have presented the reconstruction of a spectrum as a linear combination of the eigenspectra of the synthetic library (Figure~\ref{fig:pca_reconstruct}). In practice, however, the reconstructed spectrum must be further post-processed to match the observational parameters ($\vt_{\rm obs}$) of the data spectrum by applying convolutions for instrumental broadening, stellar rotation broadening, and Doppler shift, and then lastly re-sampling the spectrum to the exact pixels of the CCD (Section~\ref{subsec:postprocess}). Fortunately, because convolution is a linear operation, we can first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the reconstructed spectrum as a linear combination of these modified eigenspectra without loss of information. Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them as linear. However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the error across all pixels is smaller than 0.2\%, and thus any effect is negligible. We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra. This is the model. Then, we can rewrite Eqn~\ref{eqn:reconstruct2} as 
\begin{equation}
  f_\lambda(\vt_\textrm{obs}) \approx \widetilde{\xi}_\mu + \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}} \mathbf{w}
\end{equation}
accompanied by the conditional probability on $\mathbf{w}$ (Eqn~\ref{eqn:conditional_gp}).

The final task of designing the spectral emulator is to combine the posterior distribution of the data (Eqn~\ref{eqn:post}) with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:conditional_gp}) to form the joint distribution and then marginalize out the weights, such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.

Cite Bishop, Gelman.

The final modified covariance matrix looks like Fig A.3.

>>>>

The procedure outlined above can be a significant computational challenge.  A
typical data spectrum has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the
many repeated evaluations of the matrix product $\vR^{\trans} \vC^{-1} \vR$ in
the likelihood calculation can be numerically expensive.  However, because
$\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky
factorization to optimize the evaluation of the matrix product ($\vR^{\trans}
\vC^{-1} \vR$) and avoid the direct calculation of the matrix inverse
($\vC^{-1}$) itself.  For echelle spectra with multiple orders or multiple
spectra of the same object, taken with different instruments, the nuisance
parameters for each chunk of spectrum are independent from the other chunks. This
means that the computationally intensive steps of generating a model spectrum
for a specific wavelength range and evaluating the likelihood can be
parallelized. The only segment of the code that needs to be synchronized is the
MCMC proposal of stellar parameters, which are shared between all chunks of the
spectrum. Because all of the computationally expensive steps can be
parallelized, this algorithm can be massively parallelized on a cluster,
enabling simultaneous inference for wide spectral ranges at high resolution or
multiple spectral datasets acquired with different instruments. To sample this
model, we extend the Metropolis-Hastings sampler included in the {\tt emcee}
package \citep{foreman-mackey13} to function within a parallelized blocked
Gibbs sampler.\footnote{\url{https://github.com/iancze/emcee}}\\

@@

Alternatively, this kind of feedback could be used to make data-driven modifications to the already 
existing models, creating a new library of semi-empirical spectral models.  This could be 
accomplished by linking the parameters of the covariance kernels derived from fitting many spectra 
in a hierarchical Bayesian model, which would add confidence to the assessment that certain 
spectral features are {\it systematic} outliers and offer general quantitative guidance on how to 
weight them in the likelihood calculation.  Rather than simply assembling an empirical spectral 
library using only observations, this combined machine-learning approach would naturally provide a 
physical anchoring for the key physical parameters, since they are reflected in the spectra based 
on the physical assumptions in the original models.  This kind of large-scale analysis holds great 
promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
like those being collected in programs like the APOGEE and HERMES surveys \citep{nidever12, zucker12}), since they provide enormous leverage for identifying and improving the underlying model systematics. \\

@@

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}. We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}. Prior to the PCA decomposition, we first identify a subset of $M$ synthetic spectra which will be relevant to the star at hand. We then standardize the synthetic library by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid. The mean spectrum is 
\begin{equation}
  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
\end{equation}
$\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of stellar parameters sampled in the spectral library, while $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection. All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
\begin{equation}
  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}). We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$). We denote a single $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$. As an example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.

@@

We model each eigenspectrum weight independently; therefore there is a separate Gaussian process ($\mathcal{GP}$) for each eigenspectrum. For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$. The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
\begin{equation}
  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( 0, \mathcal{K}_\textrm{SE} (\{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})\right )
\end{equation}
The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$. Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
\begin{multline}
  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right ) \times \\
  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
\end{multline}
This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals. Then the $M$ by $M$ covariance matrix is 
\begin{equation}
\mathsf{C}^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | a_\textrm{SE}, l_\textrm{SE})
\end{equation}
which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters.

@@

@@

Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
\begin{equation}
  \begin{bmatrix}
    \mathbf{w}_k^\textrm{grid}\\
    w_k
  \end{bmatrix} \sim
  \mathcal{N} \left (
  \mathbf{0},
  \begin{bmatrix}
    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
\mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
  \end{bmatrix}
  \right )
\end{equation}
where the dependence of the covariance kernel on the optimized hyperparameters is implied. Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$. And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$. The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$). The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
\begin{equation}
  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) =  \mathcal{N} \left ( w_k \Bigl | \Bigr . \,  \mu_k , \sigma_k \right )
 \label{eqn:conditional_gp}
\end{equation}
where
\begin{equation}
 \mu_k =   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}
\end{equation}
\begin{multline}
  \sigma_k = \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \\
  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)
\end{multline}

Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).

@@

Next, we consider all $m$ of the Gaussian processes describing the weight for each eigenspectra collectively. For a given value of $\vt_\ast$, these Gaussian processes deliver a collection of $m$ one-dimensional Gaussian probability distributions. Because we have modelled each $w_k$ independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a multidimensional Gaussian with only diagonal elements filled
\begin{equation}
  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | \right . \mathbf{\mu}_\mathbf{w}, \mathbf{\Sigma}_\mathbf{w} \right )
\end{equation}
\begin{equation}
  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | 
  \begin{bmatrix}
    \mu_1 \\
    \vdots \\
    \mu_m \\
  \end{bmatrix},
  \begin{bmatrix}
    \sigma_1^2 & \ldots & 0 \\
    \vdots & \ddots & \vdots \\
    0 & \ldots & \sigma_m^2 \\
  \end{bmatrix}
  \right .
  \right)
  \label{eqn:weight_conditional}
\end{equation}

So far, we have presented the reconstruction of a spectrum as a linear combination of the eigenspectra of the synthetic library (Figure~\ref{fig:pca_reconstruct}). In practice, however, the reconstructed spectrum must be further post-processed to match the observational parameters ($\vt_{\rm obs}$) of the data spectrum by applying convolutions for instrumental broadening, stellar rotation broadening, and Doppler shift, and then lastly re-sampling the spectrum to the exact pixels of the CCD (Section~\ref{subsec:postprocess}). Fortunately, because convolution is a linear operation, we can first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the reconstructed spectrum as a linear combination of these modified eigenspectra without loss of information. Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them as linear. However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the error across all pixels is smaller than 0.2\%, and thus any effect is negligible. We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra according to the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance parameters $(\phi_\mathsf{P})$. Now, the model spectrum is specified conditional on the vector of eigenspectra weights as 
\begin{equation}
  \mathsf{M} \bigl | \bigr . \,\mathbf{w} = \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w}
\end{equation}
where 
\begin{equation}
  \mathbf{X} = \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}}
\end{equation}
Because the Gaussian process describes a probability distribution of the weights, we really have a probability distribution of possible (interpolated) models
\begin{equation}
  p(\mathsf{M} |\, \mathbf{w}) = p( \mathbf{w} |\, \vt_\ast)
\end{equation}
and the likelihood function (Eqn~\ref{eqn:likelihood}) is now also specified conditional on the weights
\begin{equation}
  p( \mathsf{D} |\, \mathsf{M}) = p( \mathsf{D} |\, \mathbf{w}) = 
  \mathcal{N} \left ( \mathsf{D} |\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w} , \mathsf{C} \right )
  \label{eqn:likelihood_conditional}
\end{equation}

The final task of designing the spectral emulator is to combine this data likelihood function with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:weight_conditional}) and then marginalize over the weights
\begin{equation}
  p(\mathsf{D} | \vt_\ast) = \int p(\mathsf{D} | \mathbf{w}) p( \mathbf{w} | \vt_\ast) d \mathbf{w}
  \label{eqn:marginal}
\end{equation}
such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model. To perform this multidimensional integral, we use a convenient lemma found in \citet[Appendix A]{gelman13}.

If the probability distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in Eqns~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the marginal distribution (Eqn~\ref{eqn:marginal}) is
\begin{equation}
  p(\mathsf{D} | \vt_\ast, \vt_\textrm{obs}, \mathbf{\Phi}) = \mathcal{N} \left ( \mathsf{D} \bigl | \bigr .\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}, \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C} \right)
\end{equation}
where the dependence on the model parameters is now made explicit.
We can couch this in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
\begin{equation}
  \mathsf{R}^\prime = \mathsf{D} - (\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w})
\end{equation}
 and
\begin{equation}
  \mathsf{C}^\prime = \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C}
\end{equation}
where $\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}$ can be thought of as the ``mean model spectrum'' given the model parameters, and the covariance matrix has been modified to account for the various possible interpolations.
